---
output: html_document
editor_options: 
  chunk_output_type: console
---




# Set up
```{r Packages}
# To clear your environment 
# remove(list = ls())

library(readxl)
library(XLConnect)
library(xlsx)
library(tidyverse)
library(dplyr)
library(proj4)
library(stringr)
library(scales)
library(lubridate)

## data describe 
library(summarytools)

library(sf)
library(maps)
library(mapdata)

library(RColorBrewer)
library(viridis)
# devtools::install_github("jaredhuling/jcolors")
library(jcolors)
```


```{r Dirs}
### Set work dir ----------------------------------------------------------
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
dirname(dir)        ## go to parent dir
setwd(dirname(dir)) ## set this parent dir as root dir
getwd()

### the data dir ----------------------------------------------------------
dir.path <- './data/DO/DO_GOM/DO_Integrated_Matli_NCSU'
setwd(dir.path)
getwd()

dir.root <- dirname(dir)
dir.fig    <- paste0(dirname(dir), '/figures/'); dir.fig
dir.output <- './data_cleaned/'

today <- format(Sys.time(), "%Y%m%d"); print(today)

### keep more decimals 
options(digits = 15)
options(pillar.sigfig = 15)
```



# Data
## Load cleaned data
```{r}
list.files(path = dir.output, pattern = 'DO_Integrated.*\\.RData', full.names = T) 
fname <- "./data_cleaned/DO_Integrated_Matli_NCSU_cleaned.RData"
load(fname)

summary(df)
dfSummary(df)

### data with DO info ------------------------------------------------------------
df <- df %>%
  dplyr::mutate(Source = as.character(Source),
                DO_na = ifelse(DO < 0, NA, DO)) %>%
  arrange(YEID, Year, Depth, DO)

df_na <- df %>%
  arrange(!is.na(DO_na))

df <- df %>%
  dplyr::filter(!is.na(DO_na),
                !is.na(Date),
                !is.na(Source)) %>% ## remove NA
  dplyr::select(-DO_na)


### data for sampling locations only ----------------------------------------------
sp <- sp %>%
  dplyr::mutate(Source = as.character(Source))

length(unique(df$YEID))  ## 8827 --> 7779
length(unique(sp$YEID))  ## 7879 --> 8117
```




## Data check
```{r}
## data description
print(paste0('Total sample points:   ', nrow(df)))
print(paste0('Total sample location: ', nrow(sp)))


### how many sampling locations from 2000-Now
sp %>%
  dplyr::filter(Year >= 2000) %>%
  distinct(YEID, .keep_all = T) %>%
  nrow()

### how many sampling locations before 2000
sp %>%
  dplyr::filter(Year < 2000) %>%
  distinct(YEID, .keep_all = T) %>%
  nrow()

## change levels of the data sources
levels(df$Source)
unique(df$Source)
mylevel <- c("SEAMAP", "LUMCON", "TAMU", "UMCES", "EPA", "NECOP", "LDWF", "latx")
df$Source <- factor(df$Source, levels = mylevel)
levels(df$Source)
## reverse the levels
df$Source <- fct_rev(df$Source)
```



```{r # of samples by yr by source}
### to see how many **samples** for each year and by sources. 
df %>%
  ungroup() %>%
  group_by(Year, Source) %>%
  tally() %>%
  ggplot() +
  geom_col(aes(x = Year, 
               y = n/1000,
               # y = n,
               fill = Source)) +
  theme_bw() + 
  
  # scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
  #             labels = trans_format("log10", math_format(10^.x))) +
  # ylab('Number of samples') +
  
  scale_x_continuous(breaks = seq(min(sp$Year), max(sp$Year), by = 1))+
  ylab(expression(paste('Number of samples (', 10^3, ')'))) +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5), 
        panel.grid.minor = element_blank())
```



```{r # of sample locations by yr and source}
### to see how many **sample locations** for each year and by sources. 
df %>%
  ungroup() %>%
  distinct(YEID, .keep_all = T) %>%
  group_by(Year, Source) %>%
  tally() %>%
  ungroup() %>%
  group_by(Year) %>%
  dplyr::mutate(sum = sum(n)) %>%
  ggplot() +
  geom_col(aes(x = Year, y = n, fill = Source)) +
  geom_text(aes(x = Year, y = sum, label = sum), hjust = 0.5, vjust = -0.5, size = 2.5) + 
  theme_bw() + 
  # scale_x_continuous(n.breaks = 10) +
  scale_x_continuous(breaks = seq(min(sp$Year), max(sp$Year), by = 1)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5),
        legend.position = c(0.1, 0.8),
        panel.grid.minor = element_blank()) +
  ylab(expression(paste('Number of locations', ''))) 
```




```{r # of samples by month}
str(df)
### how many samples in each month
df %>%
  dplyr::mutate(month = month(Date)) %>%
  arrange(!is.na(month)) %>%
  group_by(month) %>%
  tally()
  # ggplot() +
  # geom_histogram(aes(x = month))


### YEID has unique location info? --> Yes. 
df.1 <- df %>%
  ungroup() %>%
  distinct(YEID)
df.2 <- df %>%
  ungroup() %>%
  distinct(YEID, lat, lon)
```



### check data again
```{r }
### check data from UMCES -------------------------
df_um <- df %>%
  dplyr::filter(Source == 'UMCES' & Year >= 2000) 

str(df_um)
df_um %>% ggplot() + 
  geom_bar(aes(x=factor(Year))) + 
  theme_bw()


### check data from SEAMAP -----------------------
df %>%
  ungroup() %>%
  dplyr::filter(Source == 'SEAMAP' & Year >= 2000) %>%
  group_by(Source, Year) %>%
  tally()

df %>%
  ungroup() %>%
  dplyr::filter(Source == 'SEAMAP' & Year >= 2000) %>%
  distinct(YEID, .keep_all = T) %>%
  group_by(Source, Year) %>%
  tally()

  # ggplot() + 
  # geom_bar(aes(x=factor(Year))) + 
  # # geom_text(aes(label = n), hjust = 0.5, vjust = -0.5, size = 2) +
  # theme_bw()


### check data in 2015 -----------------------------
str(df)
unique(df$Source)

df.2015 <- df %>%
  filter(Source == 'SEAMAP' &
           Year == 2015)

length(unique(df.2015$YEID)) ## 2015: 94

df.2014 <- df %>%
  dplyr::filter(Source == 'LUMCON' &
           Year == 2014)
length(unique(as.character(df.2014$YEID))) ## 2014: 86
```




## Subset 2000-Now
```{r }
### subset data 2000 ~ now --------------------------
names(df)
df_after2000 <- df %>%
  dplyr::filter(Year >= 2000)
nrow(df_after2000)
print(paste0('There are ', nrow(df_after2000), ' sample data during 2000 and ', max(df$Year)))
```



##  Add SEAMAP data
  To add additional samples from NOAA_NCEI_Hypoxia Watch to `df`.
```{r}
xlsx <- list.files(path = dir.root, pattern = 'hypoxia_watch_GOM_csv_DO_Cleaned_2001_2018.xlsx', recursive = T, full.names = T); xlsx

df_hypoxia_watch <- readxl::read_excel(path = xlsx) %>%
  rename_with(str_to_title) %>%
  dplyr::mutate(Date = Yymmdd, 
                ### to create a unique id for each location; there are duplicate ids if only use `year` and `station`
                Month = month(Date),
                rowid = row.names(.),
                YEID = paste0(Year, '_hyp_watch_', str_pad(Station, 3, pad = "0"), '_', str_pad(rowid, 4, pad = "0")), 
                DO = Oxmgl,
                Depth_dem = NA, Temp = NA, Salin = NA, Instrument = NA, ShelfW =NA, NSW = NA, E = NA, N=NA, Cruise=NA, 
                Source='SEAMAP', Corr = NA, 
                lat=Latitude, lon = Longitude) %>%
  dplyr::select(names(df_after2000))


names(df_hypoxia_watch)
names(df_after2000)


### check the new id
length(unique(df_hypoxia_watch$YEID))

df_hypoxia_watch_dup <- df_hypoxia_watch %>% 
  group_by(YEID) %>% 
  mutate(n = n())  %>%
  dplyr::filter(n > 1) %>%
  arrange(YEID)



### merge the two data sets ----------------------------------------------------------------------------------------
df_final <- rbind(df_after2000, df_hypoxia_watch) %>%
  dplyr::mutate(Date = as.Date(Date))

str(df_final)


### data from hypoxia_watch has more locations in terms of spatial coverage, but does not have detailed profile (sampling at different depth); while `df` has fewer locations and more detailed profile. Therefor, `df1_2015` has some overlap with `df2_2015` --- repeated location data. But, it seems not matter much after compare these two, as there is a slight difference in the `lat` and `lon`. There are near to each other, and can serve as additional samples. 
df1_2015 <- df_after2000 %>%
  filter(Source == 'SEAMAP' &
           Depth > 8 &
           Year == 2015)

df2_2015 <- df_hypoxia_watch %>%
  filter(Source == 'SEAMAP' &
           Year == 2015)

dfs_2015 <- df_final %>%
  filter(Source == 'SEAMAP' &
           Year == 2015) 

dfs_2015_distinct <- dfs_2015 %>%
  distinct(Date, Depth, DO, .keep_all = T)


### at the meantime, we need to add spatial info from hypoxia_watch to `sp` as well. --------------------------------
sp1 <- sp 
sp2 <- df_hypoxia_watch %>%
  dplyr::mutate(OEID = NA) %>%
  dplyr::select(names(sp1)) %>%
  distinct(YEID, .keep_all = T)

names(sp1)
names(sp2)

sp_final <- rbind(sp1, sp2) %>%
  dplyr::mutate(Date = as.Date(Date))
  
str(sp_final)
```



## Save updated data as csv
```{r }

### save as csv
fname <- paste0(dir.output, 'DO_Integrated_df_final_2000_2019.xlsx');   fname
writexl::write_xlsx(x = df_final, path = fname)

fname <- paste0(dir.output, 'DO_Integrated_sp_final_2000_2019.xlsx');   fname
writexl::write_xlsx(x = sp_final, path = fname)

fname <- paste0(dir.output, 'DO_Integrated_sf_sp_final_2000_2019.RData');   fname
save(df_final, sp_final, file = fname)
```



#---------------------------
# Use the cleaned data
## - Input data for GEE
  - 1. DO profile data
  - 2. location table for extracting spectral data in GEE
```{r}
### load cleaned data ------------------------------------------------------------
fname <- paste0(dir.output, 'DO_Integrated_sf_sp_final_2000_2019.RData');   fname
load(fname)


### 1. DO profile data --> 
df_final_do <- df_final %>%
  dplyr::select(YEID, Year, Date, DO, Depth, Temp, Salin, Source, lat, lon) %>% 
  dplyr::mutate(Date = as.Date(Date))


### 2. location table (NOT SURE WE SHOULD INCLUDE hypoxia watch DATA, perhaps NOT at this moment)
df_final_pt <- sp_final %>%
  dplyr::filter(Year >= 2000) %>%
  dplyr::select(YEID, Date, Year, lat, lon) %>%
  distinct(YEID, .keep_all = T) %>%
  dplyr::mutate(Date = as.Date(Date))%>%
  dplyr::mutate(id = row.names(.))

### how many sampling locations in each year?
df_final_pt %>%
  ungroup() %>% group_by(Year) %>% tally()
  

### point to sf
projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
df_final_pt_sf <- st_as_sf(df_final_pt, coords = c("lon", "lat"), crs = projcrs)
ggplot() + geom_sf(data = df_final_pt_sf, col = "red", size = 1) + theme_bw()


### write to file ---------------------------
fname <- paste0(dirname(dir), '/data/data_for_gee/sample_2000_2019_DO.xlsx'); fname  ### xlsx can be smaller than csv file
writexl::write_xlsx(x = df_final_do, path = fname)
# fname <- paste0(dirname(dir), '/data/data_for_gee/sample_2000_2019_DO.csv'); fname 
# write.csv(x = df_final_do, file = fname, row.names = F)

fname <- paste0(dirname(dir), '/data/data_for_gee/sample_2000_2019_locations.csv'); fname
# writexl::write_xlsx(x = df2014_pt, path = fname)
write.csv(x = df_final_pt, file = fname, row.names = F)

fname <- paste0(dirname(dir), '/data/data_for_gee/sample_2000_2019_locations.shp'); fname
st_write(obj = df_final_pt_sf, dsn = fname, delete_layer = TRUE, delete_dsn = T) # overwrites
```






### - subset 2014 data for Sam
```{r}

### sampling data for 2014 --------------------
df2014 <- df_final %>%
  dplyr::filter(Year == 2014) %>%
  dplyr::select(YEID, Year, Date, DO, Depth, Temp, Salin, Source, lat, lon) %>% 
  dplyr::mutate(Date = as.Date(Date))

# dfSummary(df2014)
names(df2014)


### sampling locations that will be used for extracting Spectral_Value from images
df2014_pt <- df2014 %>%
  distinct(YEID, Date, lat, .keep_all = T) %>%
  dplyr::select(YEID, Date, Year, lat, lon) %>%
  dplyr::mutate(id = row.names(.))

projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
df2014_sf <- st_as_sf(df2014_pt, coords = c("lon", "lat"), crs = projcrs)
ggplot() + geom_sf() + geom_sf(data = df2014_sf, col = "red", size = 1) + theme_bw()



### write to file ---------------------------
fname <- paste0(dirname(dir), '/data/data_for_sam/sample_2014_DO.xlsx'); fname
writexl::write_xlsx(x = df2014, path = fname)

fname <- paste0(dirname(dir), '/data/data_for_sam/sample_2014_locations.csv'); fname
# writexl::write_xlsx(x = df2014_pt, path = fname)
write.csv(x = df2014_pt, file = fname, row.names = F)

fname <- paste0(dirname(dir), '/data/data_for_sam/sample_2014_locations.shp'); fname
st_write(obj = df2014_sf, dsn = fname, delete_layer = TRUE) # overwrites

```





## - Map samples by year & by source
```{r}
### load DO final data - `df_final`, and unique location data - `sp_final`
fname <- paste0(dir.output, 'DO_Integrated_sf_sp_final_2000_2019.RData');   fname
load(fname)

library(sf)
library(maps)
library(mapdata)

bg_transparent <- 
  theme(
    panel.background = element_rect(fill = "transparent"),             # bg of the panel
    plot.background  = element_rect(fill = "transparent", color = NA), # bg of the plot
    # panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"),    # get rid of legend bg
    # legend.box.background = element_rect(fill = "transparent"), # get rid of legend panel bg,
    legend.box.background = element_blank()
  )

shp <- map_data('usa') ## world
head(shp)




## 1. sample number stack bar by year ------------------------------------------------------------
size_key = 0.4

sp_final %>% ggplot() + 
  geom_bar(aes(x=Year, fill = Source), position = "stack") + 
  ylab('Number of sampling locations') +
  scale_x_continuous(breaks = seq(min(sp$Year), max(sp$Year), by = 1))+
  theme_bw()+
  bg_transparent +
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5),
        # legend.key.size   = unit(size_key, "cm"),
        legend.key.height = unit(size_key, "cm"),
        legend.key.width  = unit(size_key, "cm")) +
  # guides(shape = guide_legend(override.aes = list(size = size_key)), ## change the legend key size
  #        color = guide_legend(override.aes = list(size = size_key)),
  #        fill  = guide_legend(override.aes = list(size = size_key))) +
  theme(legend.position = c(0.1, 0.8)) 

pname <- paste0(dir.fig, 'total_count_sample_bySources1.png'); pname
# ggsave(filename = pname, plot = last_plot(),  bg = "transparent",
#        width = 6, height = 6, units = 'in', dpi = 300)




## 2. sample distribution facet by year ----------------------------------------------------------
ggplot() + 
  # geom_polygon(data = shp, aes(x=long, y = lat, group = group), fill = 'gray90') + 
  # scale_fill_viridis_d(option = "viridis", alpha = 0.6)+
  geom_point(data = sp_final,  # use geom_point, instead of geom_sf
             aes(x = lon, y = lat, color = Source), alpha = 0.9, size = 1.5) +
  # scale_color_brewer(palette = "Dark2") +
  facet_wrap(.~Year)+
  theme_bw() +
  xlab('Longitude') + ylab('Latitude')+
  guides(color = guide_legend(override.aes = list(size=4))) ## enlarge the key size

pname <- paste0(dir.fig, 'map_samp_by year1', '.png'); pname
# ggsave(filename = pname, plot = last_plot(), 
#        width = 14, height = 8, units = 'in', dpi = 300)




## 3. sample distribution facet by data source ---------------------------------------------------
ggplot() + 
  geom_point(data = sp_final,  # use geom_point, instead of geom_sf
             aes(x = lon, y = lat, color = Year), alpha = 0.5) +
  facet_wrap(.~Source)+
  # scale_color_gradient2(low = 'blue', mid = 'green', high = 'red', midpoint = 2000)+
  scale_color_viridis(option = "D") +
  xlab('Longitude') + ylab('Latitude')+
  theme_bw()
pname <- paste0(dir.fig, 'map_samp_by source2', '.png'); pname
# ggsave(filename = pname, plot = last_plot(), width = 14, height = 8, units = 'in', dpi = 300)

```






## - DO at multiple depth

  do_min:       minimum DO at 3-80m depth across the profile 
  do_depth_max: DO at the max depth (3-80m) across the profile
  do_20m:       DO at  3-20m depth across the profile
  do_50m:       DO at 21-50m depth across the profile
  do_80m:       DO at 51-80m depth across the profile

```{r}
names(df_final)

### do_min ------------------------------------------------
do_min <- df_final %>%
  ungroup() %>%
  dplyr::filter(Depth >= 3, Depth <=80) %>%
  group_by(YEID) %>%
  slice(which.min(DO))
length(unique(do_min$YEID))

fname <- paste0(dirname(dir), '/data/data_for_gee/sample_2000_2019_DO_min.xlsx'); fname  ### xlsx can be smaller than csv file
writexl::write_xlsx(x = do_min, path = fname)



### do_depth_max ------------------------------------------
do_depth_max <- df_final %>%
  ungroup() %>%
  group_by(YEID) %>%
  slice(which.max(Depth))
length(unique(do_depth_max$YEID))

fname <- paste0(dirname(dir), '/data/data_for_gee/sample_2000_2019_DO_depth_max.xlsx'); fname  ### xlsx can be smaller than csv file
writexl::write_xlsx(x = do_depth_max, path = fname)


### do_20m -----------------------------------------------
do_20m <- df_final %>%
  ungroup() %>%
  dplyr::filter(Depth >= 3, Depth <=20) %>%
  group_by(YEID) %>%
  ### to see the min, avg, and sd in this depth range
  # dplyr::mutate(do_20m_min = min(DO, na.rm = T),
                # do_20m_avg = mean(DO, na.rm = T),
                # do_20m_dev = sd(DO, na.rm = T),
                # n = n()) %>%
  ### take the min DO in this depth range
  slice(which.min(DO)) %>%
  as.data.frame()
fname <- paste0(dirname(dir), '/data/data_for_gee/sample_2000_2019_DO_20m.xlsx'); fname  ### xlsx can be smaller than csv file
writexl::write_xlsx(x = do_20m, path = fname)



### do_50m -----------------------------------------------
do_50m <- df_final %>%
  ungroup() %>%
  dplyr::filter(Depth >= 21, Depth <=50) %>%
  group_by(YEID) %>%
  ### to see the min, avg, and sd in this depth range
  # dplyr::mutate(do_20m_min = min(DO, na.rm = T),
                # do_20m_avg = mean(DO, na.rm = T),
                # do_20m_dev = sd(DO, na.rm = T),
                # n = n()) %>%
  ### take the min DO in this depth range
  slice(which.min(DO)) %>%
  as.data.frame()
do_50m %>%
  ggplot() +
  geom_point(aes(x = Depth, y = DO, color = Source)) +
  theme_bw()



### do_80m -----------------------------------------------
do_80m <- df_final %>%
  ungroup() %>%
  dplyr::filter(Depth >= 51, Depth <=80) %>%
  group_by(YEID) %>%
  ### to see the min, avg, and sd in this depth range
  # dplyr::mutate(do_20m_min = min(DO, na.rm = T),
                # do_20m_avg = mean(DO, na.rm = T),
                # do_20m_dev = sd(DO, na.rm = T),
                # n = n()) %>%
  ### take the min DO in this depth range
  slice(which.min(DO)) %>%
  as.data.frame()
do_80m %>%
  ggplot() +
  geom_point(aes(x = Depth, y = DO, color = Source)) +
  theme_bw()

```



### - do_min data
```{r}
do_min %>%
  ungroup() %>%
  dplyr::filter(Year >= 2005) %>%
  group_by(Year) %>%
  tally() %>%
  ggplot(aes(x=Year, y=n, fill = n))+
  geom_col(show.legend = F) +
  geom_text(aes(label = n), hjust = 0.5, vjust = -0.5, size = 3) + 
  theme_bw() +
  # theme_classic() +
  ggtitle('do_min')+ ylab('Number of sampling locations') +
  scale_x_continuous(breaks = seq(2000, max(do_min$Year), 1)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


# dfsummary %>%
#   ungroup() %>%
#   dplyr::filter(Year >= 2000) %>%
#   group_by(Year) %>%
#   tally() %>%
#   ggplot(aes(x=Year, y=n))+
#   geom_col() +
#   geom_text(aes(label = n), hjust = 0.5, vjust = -0.5, size = 2) + 
#   theme_bw() +
#   # theme_classic() +
#   ggtitle('dfsummary')+ ylab('Number of sampling locations') +
#   scale_x_continuous(breaks = seq(2000, 2017, 1)) +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
#         # panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank())
```



