---
output: html_document
editor_options: 
  chunk_output_type: inline
---


# Set up
```{r Packages, include=FALSE}
# To clear your environment 
# remove(list = ls())

library(readxl)
library(tidyverse)
library(dplyr)
library(stringr)
library(scales)
library(lubridate)

## data describe 
library(summarytools)

library(sf)
library(proj4)
library(maps)
library(mapdata)

library(RColorBrewer)
library(viridis)
# devtools::install_github("jaredhuling/jcolors")
library(jcolors)
library(scico) ## colors
```


```{r Dirs, include=FALSE}
### Set work dir ----------------------------------------------------------
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
dirname(dir)        ## go to parent dir
# setwd(dirname(dir)) ## set this parent dir as root dir
# knitr::opts_knit$set(root.dir = dirname(dir))
getwd()

### the data dir ----------------------------------------------------------
dir.root   <- dirname(dir)
dir.do     <- paste0(dir.root, '/data/for_gee/'); dir.do
dir.fig    <- paste0(dir.root, '/figures/'); dir.fig

dir.path   <- paste0(dir.root, '/data/DO_GOM/DO_Integrated_Matli_NCSU')
dir.output <- paste0(dir.path, '/data_cleaned/'); dir.output

setwd(dir.path)
# knitr::opts_knit$set(root.dir  = dir.path)
# knitr::opts_chunk$set(root.dir = dir.path)


getwd()

today <- format(Sys.time(), "%Y%m%d"); print(today)

### keep more decimals 
options(digits = 15)
options(pillar.sigfig = 15)
```


```{r Shapefile, include=FALSE}
library(sf)
library(maps)
library(mapdata)

bg_transparent <- 
  theme(
    panel.background = element_rect(fill = "transparent"),             # bg of the panel
    plot.background  = element_rect(fill = "transparent", color = NA), # bg of the plot
    # panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"),    # get rid of legend bg
    # legend.box.background = element_rect(fill = "transparent"), # get rid of legend panel bg,
    legend.box.background = element_blank()
  )

shp <- map_data('usa') ## world
head(shp)


getwd()
gom <- sf::st_read(dsn = paste0(dir.root, '/data/shp/region_rectangle.shp'))
```


# Data import

## * All DO data for GEE
  - 1. DO profile data
  - 2. location table for extracting spectral data in GEE
  
  First, define which years will be included in this analysis. For the 2nd paper, since we plan to use MODIS data only, then starting year will be 2000. This is because Terra data is avaibale since 2000/02 and Aqua is available since 2002/07. 
  
```{r}
yr_ini <- 2000
yr_end <- 2019

### load cleaned data (`df_final`) ------------------------------------------------------------
# fname <- paste0(dir.output, 'DO_Integrated_sf_sp_final_2000_2019.RData'); fname ## updated on 2022-03-08
fname <- paste0(dir.output, 'DO_Integrated_sf_sp_final_2000_2021.RData'); fname
load(fname)
names(df_final)

### 1. DO profile data
df_final_do <- df_final %>%
  dplyr::filter(Year >= yr_ini) %>%
  dplyr::filter(Year <= yr_end) %>%
  dplyr::select(-Depth_dem, -Instrument, -ShelfW, -NSW, -E, -N, -Cruise, -Corr) %>% ## remove these variables
  # dplyr::select(YEID, Year, Date, DO, Depth, Temp, Salin, Source, lat, lon) %>% 
  dplyr::filter(Depth <= 80) %>%
  dplyr::mutate(Date = as.Date(Date))


### 2. location table (NOT SURE WE SHOULD INCLUDE "hypoxia watch" data, perhaps NOT at this moment)
df_final_pt <- sp_final %>%
  dplyr::filter(Year >= yr_ini) %>%
  dplyr::filter(Year <= yr_end) %>%
  dplyr::filter(YEID %in% unique(df_final_do$YEID)) %>% ## make sure all sample locations have DO data
  dplyr::select(YEID, Date, Year, Source, lat, lon) %>%
  distinct(YEID, .keep_all = T) %>%
  dplyr::mutate(Date = as.Date(Date))%>%
  dplyr::mutate(id = row.names(.))

### --> how many sampling locations in each year?
df_final_pt %>%
  ungroup() %>% dplyr::group_by(Year) %>% tally()

### point to sf
projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
df_final_pt_sf <- st_as_sf(df_final_pt, coords = c("lon", "lat"), crs = projcrs)


## rename the data for future use
do <- df_final_do;      
pt <- df_final_pt;      
sp <- df_final_pt_sf

cat('there are', nrow(do), 'samples')
cat('there are', unique(do$YEID)%>%length(), 'sampling locations')
cat('there are', nrow(pt), 'sampling locations')
```



  *Data check*: why `do` has less location than `pt`?
```{r echo=FALSE}
t1 <- do %>%
  distinct(YEID, .keep_all = T)

t2 <- pt %>%
  distinct(YEID, .keep_all = T)

tt <- merge(x = t1 %>% dplyr::select(YEID, Depth, DO), 
            y = t2 %>% dplyr::select(YEID, id), 
            by = "YEID", all = T) %>%
  arrange(!is.na(Depth))



df_final %>% dplyr::filter(YEID == '2000_069') ## NO DATA. 

## --> action: add `dplyr::filter(YEID %in% unique(df_final_do$YEID)) %>%` to `df_final_pt`

cat('There are', nrow(t2)-nrow(t1), 'locations without DO info.')
```




```{r Save to RData, include=FALSE}
### write to file ---------------------------
fname <- paste0(dir.do, 'sample_00_19_do.xlsx'); fname  ### xlsx can be smaller than csv file
# writexl::write_xlsx(x = do, path = fname)
# fname <- paste0(dirname(dir), '/data/data_for_gee/sample_2000_2019_DO.csv'); fname 
# write.csv(x = do, file = fname, row.names = F)

fname <- paste0(dir.do, 'sample_00_19_locations.csv'); fname
write.csv(x = pt, file = fname, row.names = F)

fname <- paste0(dir.do, 'sample_00_19_locations.shp'); fname
st_write(obj = sp, fname, delete_layer = TRUE, delete_dsn = T) # overwrites

fname <- paste0(dir.do, 'sample_00_19_do_pt_sp.RData'); fname
save(do, pt, sp, file = fname)
```



```{r !! upload to GEE}
```
  Here, we need to upload 'sample_00_19_locations.shp' to GEE for extracting pixel values. 




## * Data description

```{r - Fig 2. Samples by year & by source}
### load DO final data - "do, pt, sp"
fname <- paste0(dir.do, 'sample_00_19_do_pt_sp.RData'); fname
load(fname)


## 1. sample number stack bar by year ------------------------------------------------------------
size_key = 0.4
scaleFUN <- function(x) sprintf("%.0f", x)


# pt %>% 
#   dplyr::filter(Year >= yr_ini) %>%
#   ggplot() + 
#   geom_bar(aes(x=Year, fill = Source), position = "stack") + 
#   ylab('Number of sampling locations') +
#   scale_x_continuous(n.breaks = 10, labels=scaleFUN)+
#   theme_bw()+
#   bg_transparent +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5),
#         # legend.key.size   = unit(size_key, "cm"),
#         legend.key.height = unit(size_key, "cm"),
#         legend.key.width  = unit(size_key, "cm")) +
#   theme(legend.position = c(0.9, 0.8)) 



library(wesanderson)
### this is the same plot but with labels 
pt_s <- pt  %>%
  dplyr::filter(Year >= yr_ini) %>%
  ungroup() %>%
  dplyr::distinct(YEID, .keep_all = T) %>%
  group_by(Year, Source) %>%
  tally() %>%
  ungroup() %>%
  group_by(Year) %>%
  dplyr::mutate(sum = sum(n)) %>%
  ## remove the previous factor levels
  dplyr::mutate(Source = as.character(Source))

unique(pt_s$Source)

Source_levels <- c('LDWF', 
                   # 'NECOP', 
                   'EPA', 
                   'UMCES', 
                   'TAMU', 'LUMCON', 'SEAMAP')
Source_labels <- c(
  'Louisiana Department of Wildlife and Fisheries (LDWF)',
  # 'Nutrient Enhanced Coastal Productivity Program (NECOP)', 
  'U.S. Environmental Protection Agency (EPA)', 
  'University of Maryland Center for Environmental Sciences (UMCES)', 
  'Texas A&M University (TAMU)', 
  'Louisiana Universities Marine Consortium (LUMCON)', 
  'Southeast Area Monitoring and Assessment Program (SEAMAP)'
)

pt_s %>%
  dplyr::mutate(Source = factor(Source, levels = Source_levels, labels = Source_labels)) %>%
  ggplot() +
  geom_col(aes(x = Year, y = n, fill = Source), alpha = .7) +
  geom_text(aes(x = Year, y = sum, label = sum), hjust = 0.5, vjust = -0.5, size = 2.5) + 
  theme_bw() + 
  scale_x_continuous(#n.breaks = 19, labels=scaleFUN, 
                     breaks = seq(from = 2000, to = 2019, by = 1)) +
  scale_fill_brewer(palette = "Set2", 
                    # name = 'Field-collected \ndissolved oxygen (DO) \nsample sources',
                    name = 'Field-collected dissolved oxygen (DO) sample sources',
                    direction = -1, 
                    ) +
  # scale_fill_manual(values = wes_palette("Royal1", n = 6, type = 'continuous')) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5),
        legend.key.height = unit(size_key, "cm"),
        legend.key.width  = unit(size_key, "cm"),
        # legend.position = c(0.85, 0.8),  ## if use abbreviation only
        legend.position = "bottom",        ## if use the full names
        panel.grid.minor = element_blank()) +
  ## if use the full names
  guides(fill = guide_legend(ncol = 1, direction = "vertical")) +
  ylab(expression(paste('Number of sampling locations', ''))) 
pname <- paste0(dir.fig, 'sample_num_bySource_DZT_fullname.png'); pname
ggsave(filename = pname, plot = last_plot(), width = 7, height = 5+2, units = 'in', dpi = 300) ## bg = "transparent",
```



```{r - Fig S2. Sample map by year}
## 2. sample distribution facet by year --------------------------------------------------
pt %>% 
  dplyr::filter(Year >= yr_ini) %>%
  ggplot() + 
  geom_point(aes(x = lon, y = lat, color = Source), alpha = 0.7, size = 0.2) +
  # scale_color_brewer(palette = "Dark2") +
  facet_wrap(.~Year)+
  theme_bw() +
  xlab('Longitude') + ylab('Latitude')+
  guides(color = guide_legend(override.aes = list(size=4))) ## enlarge the key size

pname <- paste0(dir.fig, 'sample_map_byYear_R1.png'); pname
ggsave(filename = pname, plot = last_plot(), width = 7, height = 5, units = 'in', dpi = 300)
```
```{r}
pt %>% 
  dplyr::filter(Year >= yr_ini) %>%
  # dplyr::mutate(Source = fct_rev(Source)) %>%
  ggplot() + 
  geom_sf(data = gom, color = "NA", fill = "NA") + 
  geom_point(aes(x = lon, y = lat, color = Source, size = Year, alpha=-Year), 
             # alpha = 0.4, 
             shape = 1) +
  # scale_color_brewer(palette = "Dark2") +
  # scale_size_continuous(trans = ) +
  scale_size_binned(nice.breaks = T, n.breaks = 5, range = c(0,5))+
  scale_alpha_binned(range = c(0.1, 1))+
  theme_bw() +
  xlab('Longitude') + ylab('Latitude')+
  guides(color = guide_legend(override.aes = list(size=4))) + ## enlarge the key size
  theme(legend.position="bottom", 
        # legend.direction = "horizontal", 
        legend.direction = "vertical", 
        # legend.box ="vertical", 
        legend.box = "horizontal",
        # legend.spacing.x = unit(10, "mm"),
        legend.box.margin = margin(0, 0, 0, 0)) + ## legend.box="vertical", legend.margin=margin()
  guides(fill=guide_legend(nrow=2, byrow=TRUE),
         label.theme = element_text(angle = 90, vjust = 0.5))

pname <- paste0(dir.fig, 'sample_map_byYear_bySource2.png'); pname
ggsave(filename = pname, plot = last_plot(), width = 7, height = 5, units = 'in', dpi = 300)
```


```{r - Fig X. Sample map by source}
## 3. sample distribution facet by data source -------------------------------------------
pt %>% 
  ggplot() + 
  geom_point(aes(x = lon, y = lat, color = Year), alpha = 0.4, size = 0.2) +
  facet_wrap(.~Source)+
  # scale_color_gradient2(low = 'blue', mid = 'green', high = 'red', midpoint = 2000)+
  scale_color_viridis(option = "D") +
  xlab('Longitude') + ylab('Latitude')+
  theme_bw()
pname <- paste0(dir.fig, 'sample_map_bySource_update2019_R1.png'); pname
# ggsave(filename = pname, plot = last_plot(), width = 7, height = 5, units = 'in', dpi = 300)

```






## * DO bottom

  do_min:       minimum DO at 3-80m depth across the profile 
  do_bottom: DO at the max depth (3-80m) across the profile
  do_20m:       DO at  3-20m depth across the profile
  do_50m:       DO at 21-50m depth across the profile
  do_80m:       DO at 51-80m depth across the profile

```{r}
names(do)

### depths distribution --------------------------------
max(do$Depth) 

do %>% 
  # group_by(YEID, lat, lon) %>% tally() %>%
  # dplyr::filter(nchar(YEID) < 10) %>%
  as.data.frame() %>%
  ggplot()+
  # geom_violin(mapping = aes(x = 1, y = n)) + 
  # facet_wrap(~Year) +
  geom_histogram(aes(x = Depth)) +
  theme_bw()



### do_bottom ------------------------------------------
do_bottom <- do %>%
  ungroup() %>%
  group_by(YEID, Year) %>%
  slice(which.max(Depth))
length(unique(do_bottom$YEID))
length(unique(do$YEID))



s1 <- do_bottom %>% group_by(Year) %>% tally()
s2 <- do_bottom %>% dplyr::filter(nchar(YEID) < 10) %>% group_by(Year) %>% tally()
cbind(s1, s2$n)


fname <- paste0(dir.do, 'sample_2000_2019_do_bottom.xlsx'); #fname
writexl::write_xlsx(x = do_bottom, path = fname)
```



```{r data desc - temporal - Fig S3}
library("wesanderson")

### this is hard to see the trend
do_bottom %>%
  # dplyr::filter(nchar(YEID) < 10) %>%   ## ----> optional !!!!
  ggplot(aes(x = Date, y = DO)) +
  geom_point() +
  theme_bw()

### perhaps to look at several sub-regions, using fishnet?
do_bottom %>%
  # dplyr::filter(nchar(YEID) < 10) %>%
  ungroup() %>%
  distinct(YEID, Date, .keep_all = T) %>%
  dplyr::mutate(mm = month(Date)) %>%
  dplyr::filter(mm <=9) %>%
  dplyr::mutate(mm = as.factor(mm),
                mmdd = paste(2000, month(Date), day(Date), sep = '-'),
                mmdd = as.Date(mmdd)) %>%
  group_by(YEID, Date, Year, mmdd, mm) %>% tally() %>%
  as.data.frame() %>%
  ggplot()+
  geom_col(aes(x = mmdd, y = n, fill = mm)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  scale_fill_manual('Month', values = wes_palette("Zissou1")) +
  facet_wrap(~Year) + ## , scales = 'free_x'
  theme_bw() +
  # theme(axis.text.x = element_text(angle = 45, hjust = 1,vjust = 0.5)) +
  ggtitle('Number of sample locations by month') + 
  ylab('Count') + xlab("Month")
  

pname <- paste0(dir.fig, 'SM_Number of samples by date_R1.png'); pname
ggsave(filename = pname, plot = last_plot(), width = 7*1.5, height = 4*1.5, units = 'in', dpi = 300)
```


```{r data desc - temporal - SEAMAP}
### For each agency, sampling date 
levels(do_bottom$Source)


do_bottom %>%
  # dplyr::filter(nchar(YEID) < 10) %>%
  # dplyr::filter(Source == 'LUMCON') %>%
  dplyr::filter(Source == 'SEAMAP') %>%
  ungroup() %>%
  distinct(YEID, Date, .keep_all = T) %>%
  dplyr::mutate(mm = month(Date)) %>%
  dplyr::filter(mm <=9) %>%
  dplyr::mutate(mm = as.factor(mm),
                mmdd = paste(2000, month(Date), day(Date), sep = '-'),
                mmdd = as.Date(mmdd)) %>%
  group_by(YEID, Date, Year, mmdd, mm) %>% tally() %>%
  as.data.frame() %>%
  ggplot()+
  geom_col(aes(x = mmdd, y = n, fill = mm)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  scale_fill_manual('Month', values = wes_palette("Zissou1")) +
  facet_wrap(~Year) + ## , scales = 'free_x'
  theme_bw() +
  # theme(axis.text.x = element_text(angle = 45, hjust = 1,vjust = 0.5)) +
  ggtitle('Number of sample locations by month') + 
  ylab('Count') + xlab("Month")
```



```{r data desc - spatial}
### this is hard to see the trend
do_bottom %>%
  dplyr::filter(nchar(YEID) < 10) %>%
  # dplyr::filter(Year == 2014) %>%
  ggplot(aes(x = lon, y = lat, color = DO)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scico::scale_color_scico(palette = "roma") +
  facet_wrap(~Year) +
  theme_bw() +
  ggtitle('Bottom DO map') +
  theme(panel.grid.minor = element_blank(), legend.position = 'bottom')

fname <- paste0(dir.fig, 'sample_map_do_bottom.png'); fname
ggsave(filename = fname, plot = last_plot(), width = 8, height = 6, units = 'in', dpi = 300)
```

