---
title: "Untitled"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source('./code/_path.R')
```



## Percentile of global dead zone size 
```{r}
f <- './data/eutrophication_hypoxia map data Set_WRI/diaz-2008-TableS1.xlsx'

d <- readxl::read_excel(f) %>%
  mutate(Area_km2 = as.numeric(Area_km2))

summary(d$Area_km2)
numbers <- d$Area_km2

# Number to find the quantile for
target_number <- 5000

# Compute the empirical cumulative distribution function
ecdf_func <- ecdf(numbers)

# Get the quantile of the target number
target_percentile <- ecdf_func(target_number) * 100

# Print the quantile
print(target_percentile)
```


```{r}
# Create a data frame for the target point
target_df <- data.frame(
  Percentile = target_percentile,
  Value = target_number
)

# Calculate every 10th percentile
percentiles <- quantile(numbers, probs = seq(0, 1, by = 0.1), na.rm = TRUE)

# Create a data frame for plotting
df <- data.frame(
  Percentile = as.numeric(gsub("%", "", names(percentiles))),  # Convert to numeric and scale to 0-100
  Value = as.numeric(percentiles)                    # Corresponding values
)

# Plot the percentiles
ggplot(df, aes(x = Percentile, y = Value)) +
  geom_point(size = 3, color = "blue") +
  geom_line(group = 1, color = "blue") +
  geom_text(aes(label = Value), vjust = -0.5, size = 3.5) +  # Add labels above points
  # Add the target number as a special point
  geom_point(data = target_df, aes(x = Percentile, y = Value), 
             color = "red", size = 2, shape = 17, alpha = 0.7) +  # Red triangle
  geom_text(data = target_df, aes(x = Percentile, y = Value, label = Value), 
            color = "red", vjust = -0.5, size = 4) +  # Label the target point
  labs(title = "Every 10th Percentile of Numbers",
       x = "Percentile",
       y = "Value") +
  scale_x_continuous(breaks = seq(0, 100, by = 10)) +  # Ensure x-axis goes from 0 to 100
  theme_bw()
```



## Nutrient loads to the Gulf
  Casey Lee, 2022, Nutrient loads to the Gulf of Mexico produced by the USGS National Water Quality Network, 1968-2021: U.S. Geological Survey, https://doi.org/10.5066/P9G0EEUE.
  
  
  #`SITE_ABB`: A text abbreviation of the site name primarily for use within the National Water Quality Program											
  
  									
  `CONSTIT`: A character abbreviation indicative of the water-quality constituent being reported. Values include 
  NO3_NO2: nitrate plus nitrite filtered as N; 
  *TN: total nitrogen as N*; 
  *TP: total phosphorus as P*; 
  SSC: suspended-sediment concentration; 
  NH3: ammonia filtered as N; 
  OP: orthophosphate filtered	 as P; 
  SI: silica filtered as SiO2; 
  CHL: Filtered chloride; 
  TDS: Total Dissolved Solids	
  
  `WY`: A numeric identifier of the water year in which data were recorded.  A water year is defined as the 12-month period from October 1 for any given year through September 30 of the following year. The water year is designated by the calendar year in which it ends.											
  
  `MODTYPE`: A character indicator of the type of model used to compute water-quality loads as indicated in Lee and others (2017). Values include
  WRTDS_K indicating the WRTDS-K method; 
  WRTDS_FNG indicating flow-normalized loads computed using the WRTDS generalized flow-normalization method; and 
  WRTDS_FNS indicating flow-normalized loads computed using the WRTDS stationary flow-normalization method	
  
  `TONS`: *Annual water-quality or flow-normalized annual water-quality load measured in tons*											
  `TONS_L90`: Lower 90 percent confidence interval about the annual or flow-normalized annual water-quality load measured in tons. NA values indicate that confidence intervals were not computed.											
  `TONS_U90`: Upper 90 percent confidence interval about the annual or flow-normalized annual water-quality load measured in tons. NA values indicate that confidence intervals were not computed.	
  
  `CONC`: *Annual flow-normalized concentration measured in milligrams per liter*. Values are only published along with annual flow-normalized water-quality loads	 NA values are published along with WRTDS-K loads)										
  `CONC_L90`: Lower 90 percent confidence interval about the annual flow-normalized water-quality load measured in tons. NA values indicate that confidence intervals were not computed.											
  `CONC_U90`: Upper 90 percent confidence interval about the annual flow-normalized water-quality load measured in tons. NA values indicate that confidence intervals were not computed.											

```{r}
f <- './data/Casey Lee, 2022, Nutrient loads to the Gulf of Mexico/WRTDSloads.csv'

d <- readr::read_csv(f, skip = 14, show_col_types = F)

names(d)
unique(d$CONSTIT)
unique(d$MODTYPE)

var <- 'TN'
# var <- 'TP'

if(var == 'TN'){
  var_title <- 'Total nitrogen'
} else if(var == 'TP') {
  var_title <- 'Total phosphorus'
} else {
  warning('Please provide more information!')
}

d_tn <- d %>%
  filter(SITE_ABB == 'GULF',
         CONSTIT == var,
         MODTYPE == 'WRTDS_FNS',
         WY >= 1990 & WY <= 2020)

p1 <- d_tn %>%
  ggplot(aes(x = WY, y = TONS)) +
  geom_point(size = 3, color = "blue") +
  geom_line(group = 1, color = "blue") +
  geom_vline(xintercept = 2000, linetype = 'dashed', color = 'red', alpha = .5) + 
  # facet_wrap(~MODTYPE) +
  theme_bw() +
  ggtitle(paste0(var_title, ' load \nmeasured at the Gulf site (tons)')) +
  xlab('Water year') + ylab('Load (tons)') +
  labs(caption = "")

p2 <- d_tn %>%
  ggplot(aes(x = WY, y = CONC)) +
  geom_point(size = 3, color = "blue") +
  geom_line(group = 1, color = "blue") +
  geom_vline(xintercept = 2000, linetype = 'dashed', colour = 'red', alpha = .5) + 
  # facet_wrap(~MODTYPE) +
  theme_bw() +
  ggtitle(paste0(var_title, ' concentration \nmeasured at the Gulf site (mg/l)')) +
  xlab('Water year') + ylab('Concentration (mg/l)') +
  labs(caption = "Data source: Casey Lee, 2022, USGS (https://doi.org/10.5066/P9G0EEUE)")


library(ggpubr)
ggarrange(p1, p2)

fname <- paste0(dir.fig, 'SM_', var, ' loads to the Gulf.png'); fname
ggsave(filename = fname, plot = last_plot(), width = 8, height = 5, units = 'in', dpi = 300)
```
  
