---
title: "Figure2"
author: "Yingjie"
date: "10/7/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---


  This script aims to combine sub-plots for Figure 2. 



# Setup 
```{r}
remove(list = ls())

getwd()

library(tidyverse)
library(dplyr)
library(Rmisc) ## `summarySE()` provides standard deviation, standard error of the mean, and a (default 95%) confidence interval

library(ggplot2)
library(ggpubr)
library(viridis)
library(tidyquant)

```



# Data


## Load data from 3 models. 
  
```{r}

# prb <- 0       # to get the actual range (min to max)
prb <- 2.5/100 # to get 95% quantiles range
# prb <- 5/100   # to get 90% quantiles range


## 2. load RF model error data -----------------------------------------------------------
list.files(path = './data/results_RF/', pattern = '^rf_r2_mse_mae_', full.names = T)

fname <- paste0('./data/results_RF/rf_r2_mse_mae_by_yearNO.csv');         ## Zilong
fname <- paste0('./data/results_RF/rf_r2_mse_mae_by_yearNO_12vars.csv');  ## + doy_img
err_rf <- readr::read_csv(fname)



## for the testing set -------------------------------------------------------------------
acc_set <- 'testing'

### input data for plotting 
err_rf_test <- err_rf %>%
  dplyr::rename(RMSE = mse) %>%
  gather(key = 'errType', value = 'value', r2:RMSE) %>%
  dplyr::mutate(errType = toupper(errType),
                errType = factor(errType, levels = c('R2', 'RMSE', 'MAE')))

### option 1. use mean, se OR sd,for plotting ----

# rfErr_test <- summarySE(data = err_rf_test, measurevar="value", groupvars=c("year", "nw", "errType")) %>%
#   dplyr::select(-year, -N) %>%
#   dplyr::rename(lag = nw)



### option 2. use median, max, min for plotting -----
rfErr_test <- err_rf_test %>%
  dplyr::select(-year) %>%
  group_by(lag, errType) %>%
  slice_sample(n = 20) %>%
  dplyr::summarise(med = median(value, na.rm = T),
                   # max = max(value, na.rm = T), 
                   # min = min(value, na.rm = T), 
                   min = quantile(value, prb),
                   max = quantile(value, probs = 1-prb),
                   avg = mean(value, na.rm = T),
                   iqr = IQR(value, na.rm = T),       ## IQR = Q3-Q1 ('Interquartile Range' = 'third quartile' - 'first quartile')
                   lci = CI(value, ci = 0.95)['lower'],
                   uci = CI(value, ci = 0.95)['upper']
                   ) %>% 
  # dplyr::select(-avg, -iqr) %>% ## updated on 4/21/2022 using new data
  dplyr::mutate(trainingSampErr = NA) %>%
  dplyr::filter(lag <= 80) %>%
  as.data.frame()
  



## for the training set ------------------------------------------------------------------
## --> add within-sample-error (i.e., training accuracy) of RF model
# rfErr_train <- err_rf_train %>%
#   gather(key = 'errType', value = 'value', RMSE:R2) %>%
#   dplyr::mutate(errType = factor(errType, levels = c('R2', 'RMSE', 'MAE'))) %>%
#   dplyr::select(-year) %>%
#   dplyr::rename(lag = nw) %>%
#   group_by(lag, errType) %>%
#   dplyr::summarise(med = median(value, na.rm = T), ## median
#                    # max = max(value, na.rm = T), 
#                    # min = min(value, na.rm = T), 
#                    min = quantile(value, prb),
#                    max = quantile(value, probs = 1-prb),
#                    avg = mean(value, na.rm = T),
#                    sd  = sd(value, na.rm = T),
#                    iqr = IQR(value, na.rm = T)) %>%
#   # dplyr::select(-avg, -iqr) %>%
#   dplyr::filter(lag <=80) %>%
#   as.data.frame()
# 
# ## --> final RF 
# rfErr_test$trainingSampErr <- rfErr_train$avg





## 4. get the complete_rate of RS band 

# complete_rate <- err_rf %>%
#   distinct(year, nw, complete_rate_avg) %>%
#   dplyr::mutate(nw = - nw) %>%
#   dplyr::rename(lag = nw)
# 
# mean(complete_rate$complete_rate_avg) ## 0.35
# 
# complete_rate %>%
#   dplyr::filter(lag >= -80) %>%
#   ggplot() +
#   geom_line(aes(x = lag, y = complete_rate_avg), color = 'orange') +
#   theme_bw()
```





## Table 1 - result summary

  This table reports the accuracy information based on best time windows (or time lag) we choose for model inputs. 
  
  For "Training error", we report the `median` values for each model and each indicator;
  
  For "Testing error",  we *additionally* report the `lower and upper percentiles`. 

```{r eval=FALSE, include=FALSE}
names(rfErr_test)


### for training data set ----------------------------------------------------------------

### --> 1. if look at the most recent 4 days
# rfErr_train %>%
#   ungroup() %>%
#   dplyr::filter(lag < 5) %>%
#   group_by(errType) %>%
#   # dplyr::summarise() %>%
#   arrange(errType, desc(med)) %>%
#   as.data.frame()


### --> 2. if look at the whole 80 days range
# err_rf_train %>%
#   gather(key = 'errType', value = 'value', RMSE:R2) %>%
#   dplyr::mutate(errType = factor(errType, levels = c('R2', 'RMSE', 'MAE'))) %>%
#   dplyr::select(-year) %>%
#   dplyr::rename(lag = nw) %>%
#   # dplyr::filter(lag < 5) %>%
#   dplyr::filter(lag < 5) %>%
#   group_by(errType) %>%
#   dplyr::summarise(
#     # med = median(value, na.rm = T), ## median
#                    # max = max(value, na.rm = T), 
#                    # min = min(value, na.rm = T), 
#                    # min = quantile(value, prb),
#                    # max = quantile(value, probs = 1-prb),
#                    avg = mean(value, na.rm = T),
#                    sd = sd(value, na.rm = T),
#                    iqr = IQR(value, na.rm = T)) %>%
#   as.data.frame()








### for testing data set -----------------------------------------------------------------
# rfErr_test %>% 
#   ungroup() %>%
#   dplyr::filter(lag < 5) %>%
#   group_by(errType) %>%
#   # dplyr::summarise() %>%
#   arrange(errType, desc(med)) %>%
#   as.data.frame()

library(plotrix)

err_rf_test %>%
  dplyr::ungroup() %>%
  dplyr::select(-year) %>%
  dplyr::ungroup() %>%
  # dplyr::filter(lag < 5) %>%       ## what time window to use!!!
  dplyr::filter(lag < 33 & lag > 29) %>%
  dplyr::group_by(errType) %>%
  dplyr::summarise(
    # med = median(value, na.rm = T),
    # q1 = quantile(value, probs = 1/4) %>% as.numeric(),
    # q3 = quantile(value, probs = 3/4) %>% as.numeric(),
    # d1 = q1 - med,
    # d3 = q3 - med,
    avg = mean(value, na.rm = T),
    sd = sd(value, na.rm = T),
    # n  = n(),
    # lse = STDERR(value)['lower'], ## lower standard error
    # lse = STDERR(value)['upper'],
    # mse = STDERR(value)['mean'],
    # se2 = std.error(value),
    # lse2 = mse + se2,
    lci = CI(value, ci = 0.9)['lower'],
    uci = CI(value, ci = 0.9)['upper']
                   
                   ) %>% 
  dplyr::mutate(trainingSampErr = NA) %>%
  as.data.frame()


```



## Plot 3 models 

```{r}
### 1. options to plot lines -------------------------------------------------------------
# line_to_be_plot <- "med" ## use `median` values to plot the lines
line_to_be_plot <- "avg"   ## use `mean`


### 2. options to plot shade -------------------------------------------------------------
# upper = 'max'; lower = 'min'
upper = 'uci'; lower = 'lci'



### 3. range of lag to plot --------------------------------------------------------------
lag_range <- 30
lag_range <- 80
# lag_range <- 60


### 4. if to plot the complete_rate of data 
# complete_rate_add <- T
complete_rate_add <- F
# complete_rate_dat <- complete_rate %>% dplyr::filter(lag >= -lag_range)

if(complete_rate_add == T) {
  complete_rate_alpha <- 1
} else {
  complete_rate_alpha <- 0
}
                  



### 5. 
if_smooth <- F
# if_smooth <- T

bandwidth_smooth <- ifelse(if_smooth == T, 5, NA)




### plot ---------------------------------------------------------------------------------
model_fullname <- c('Random Forest Regression')
model_levels   <- c('RFR')
err_levels     <- c('R2', 'RMSE', 'MAE')

## 1. the input data for plot
err_3models <- rbind(
                     rfErr_test %>% dplyr::mutate(mod = 'RFR')) %>%
  dplyr::mutate(lag = -lag) %>%
  dplyr::rename('value' = all_of(line_to_be_plot)) %>%
  dplyr::rename('upper' = all_of(upper)) %>%
  dplyr::rename('lower' = all_of(lower)) %>%
  
  dplyr::filter(lag >= -lag_range) %>%
  
  group_by(mod, errType) %>%
  # dplyr::mutate(med_avg = mean(value, rm.na = T)) %>% ## to get the mean of mean (or median) value
  ungroup() %>%
  arrange(mod, errType) %>%
  dplyr::mutate(mod     = factor(mod,     levels = model_levels, labels = model_fullname),
                errType = factor(errType, levels = err_levels)) %>%
  dplyr::select(lag, errType, value, everything()) %>%
  as.data.frame()


str(err_3models)

## 2. identify the best day by looking for the date with the highest model accuracy 
# err_3models_bestday <- err_3models %>% 
#   dplyr::mutate(mod = as.factor(mod), errType = as.factor(errType)) %>%
#   group_by(mod, errType) %>%
#   dplyr::mutate(best    = ifelse(errType != 'R2', min(med), max(med, na.rm = T)),
#                 bestday = ifelse(med == best, lag, NA)) %>%
#   arrange(mod, errType, med) %>%
#   # dplyr::filter(!is.na(bestday)) %>%
#   as.data.frame()




### smooth the results -------------------------------------------------------------------
# err_ls <- unique(err_3models$errType) %>% as.character(); err_ls
# mod_ls <- unique(err_3models$mod) %>% as.character()
# 
# err_3models_smooth <- data.frame()
# 
# 
# for (m in mod_ls) {
#   print(m)
#   
#   for (e in err_ls) {
#     print(e)
#   
#     df0 <- err_3models %>% 
#       ungroup() %>%
#       as.data.frame() %>%
#       # dplyr::mutate(mod = as.character(mod)) %>%
#       dplyr::filter(mod == m) %>%
#       dplyr::filter(errType == e)
#     
#     df1 <- ksmooth(x = df0$lag, y = df0$value, 'normal', bandwidth=bandwidth_smooth, n = nrow(df0)) %>% 
#       as.data.frame() %>%
#       dplyr::rename(lag = x, value = y) %>%
#       left_join(x = df0 %>% dplyr::select(-value), 
#                 y = ., 
#                 by = 'lag') 
#     
#     err_3models_smooth <- rbind(err_3models_smooth, df1)
# }
# }
# 
# 
if (if_smooth == T) {
  err_3models_input <- err_3models_smooth
} else {
  err_3models_input <- err_3models %>% as.data.frame()
}








### 3. plot ------------------------------------------------------------------------------

lag_max <- max(err_3models_input$lag, na.rm = T); lag_max
lag_min <- min(err_3models_input$lag, na.rm = T); lag_min

str(err_3models_input)


(p <- 
    ggplot(data = err_3models_input %>% dplyr::rename(Training = trainingSampErr), 
           aes(x=lag, 
               y=value,
               # color = errType,
               # fill = errType
               ), 
           size = 0.5)+
   
    # geom_line()+
    geom_point(shape = 16, size = .8, color = 'gray40', alpha = 0.6) +
    
    # geom_line(aes(y=Training, color = 'Training'), 
    #           # col='red', 
    #           size  = 0.4,
    #           alpha = 0.4, linetype = 'solid', show.legend = T) + ## 'dashed', 'dotted'
    
    
   #  labs(colour="Datasets",x="xxx",y="yyy")+
   # scale_color_manual(values = c("blue","red"))+
   # scale_color_viridis() +
    # geom_line(data = complete_rate_dat, aes(x = lag, y = complete_rate_avg), color = 'orange', alpha = complete_rate_alpha) +
    # geom_ribbon(aes(ymax = upper, ymin = lower), alpha = 0.2)+
    geom_smooth(method = 'loess', 
                span = 0.1,
                formula = 'y~x', linetype=1, 
                color = 'gray40',
                alpha = .3,
                size = 0.4) +
    
    scale_x_continuous(breaks = seq(lag_min, lag_max, 10), limits = c(lag_min, lag_max)) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
    
    ## --> To highlight the best data range for prediction  
    # geom_vline(data = err_3models_bestday,  aes(xintercept = bestday), linetype = "dashed", color="blue") + 
   geom_vline(xintercept = -31, linetype = 'dashed', alpha = 0.3) +
  
    
   # geom_rect(
   #    # data = data.frame(mod = factor(c('Lagged Linear Regression'))), ## not to add 'rect' for FDA
   #    data = data.frame(mod = factor(model_fullname[1:3])), ## not to add 'rect' for FDA
   #    aes(x = NULL, y = NULL, xmin=lag_min,xmax=-30, ymin=-Inf, ymax=Inf),
   #    fill = 'gray', alpha=0.2, show.legend = F) +
    geom_rect(
      data = data.frame(mod = factor(c('Random Forest Regression'))), ## not to add 'rect' for FDA
      aes(x = NULL, y = NULL, xmin=-32,xmax=-30, ymin=-Inf, ymax=Inf),
      fill = 'blue', alpha=0.1) +
   #  geom_rect(
   #    data = data.frame(mod = factor(c('Lagged Linear Regression', 'Random Forest Regression'))), ## not to add 'rect' for FDA
   #    aes(x = NULL, y = NULL, xmin=-19,xmax=-15, ymin=-Inf, ymax=Inf), 
   #    fill = 'blue', alpha=0.1, show.legend = F) + 
   
   # geom_hline(yintercept = 0.6, linetype = 'dashed', alpha = 0.3) +
  
  
  
  
    facet_grid(errType~mod, scales = 'free_y', switch = 'y')+
    labs(x='Time lag', y='Model Accuracy') +
    # 
    #  scale_colour_manual(name = "Legend",
    #                      values = c("Testing" = "black", "Training"= "red")) +
   
   
    theme_bw() +
    theme(
          # legend.position = c(0.9, 0.74), ## right side: 0.92, 0.74
          legend.position="none",
          legend.key=element_blank(),  ## remove the key background
          legend.key.height = unit(0.01, 'npc'),
          legend.title = element_blank(), 
          legend.background = element_rect(fill='transparent'),
          legend.box.background = element_rect(fill='transparent', color = 'transparent'), 
          panel.grid.minor = element_blank())
)
# library(lemon)
# reposition_legend(p, 'left')
# ggsave('./figures/accuracy_comparison_3models80d.png', p, width=7, height=5)


pname <- paste0('./figures/accuracy_', 
                '95CI_lagRange', lag_range, '_', gsub('rf_r2_mse_mae_|.csv', '', basename(fname)),
                '.png'); pname
ggsave(filename = pname, p, width=7, height=5)
```








