---
title: "Untitled"
author: "Yingjie"
date: "11/1/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---


Tidy the data table extracted from MODIS images
YingjieLi.edu@gmail.com
update on: 2019-11-01


# Setup
```{r Dirs and Packages}
### To clear your environment 
# remove(list = ls())

# set work dir
path <- rstudioapi::getSourceEditorContext()$path
dir  <- dirname(rstudioapi::getSourceEditorContext()$path); dir
setwd(dir)
getwd()
today <- format(Sys.time(), "%Y%m%d"); print(today)

library(tidyverse)
library(dplyr)
library(rlist)
library(lubridate)
library(cowplot)
```


```{r Functions}
### define a convert function 
ms_to_date = function(ms, t0="1970-01-01", timezone) {
        ## @ms: a numeric vector of milliseconds (big integers of 13 digits)
        ## @t0: a string of the format "yyyy-mm-dd", specifying the date that
        ##      corresponds to 0 millisecond
        ## @timezone: a string specifying a timezone that can be recognized by R
        ## return: a POSIXct vector representing calendar dates and times        
        sec = as.numeric(ms) / 1000
        as.POSIXct(sec, origin=t0, tz=timezone)
}
```





###################################### #
#  Data
###################################### #

##  Station info (st)
```{r}

```


##  RS data
  test code
```{r eval=FALSE, include=FALSE}
path_data <- './Img2Table/Img2Table_04_20191115'
sensor    <- 'aqua_'
# sensor    <- 'terr_'
yr        <- 2015

# read csv
pattern_file <- paste0('^', year, '_', sensor, '*'); print(pattern_file)
csv_ls <- list.files(path = path_data, pattern = pattern_file, full.names = T)
csv_ls; length(csv_ls) ## should be 14

## a blank df
data <- data.frame()

for (csv in csv_ls) {
  # print(csv)
  band_name <- gsub(pattern = '.csv', replacement = '', 
                    x = gsub(pattern = paste0(path_data, '/'), 
                             replacement = '', x = csv))
  print(band_name)
  
  csv_df <- read.csv(file = csv, header = T, stringsAsFactors = F)
  csv_df <- cbind(band = band_name, csv_df) ## add band name col
  data <- rbind(data, csv_df)
}

## check data lenght (total row number)
print(length(data$system.index)) ## final combined data
print(length(csv_ls) * length(csv_df$system.index)) ## 14 csv files * 243 stations in each 

### change the col names
### update the coverage of data on July 9, 2019

## way#1
# date_list <- names(data)[3:(length(names(data))-21)]; date_list
# date_list <- gsub('X', '',  date_list); date_list
# date_list <- gsub('\\.', '-',  date_list); date_list

## way#2
date_list <- seq(from = as.Date(paste0(yr, '-02-01')), 
                 to   = as.Date(paste0(yr, '-09-30')), by = 1) %>% as.character()
# cols_dates <- gsub('\.', '/', cols_dates)
print(date_list) ## 107 ## 241
names(data)[3:(3+length(date_list)-1)] <- as.character(date_list)
names(data)

## subset data
data2 <- data %>% 
  dplyr::select(
    band, date_list, 
    matches('dateutc|station|latitude|longitude|oxmgl|depth', ignore.case = T))
names(data2)

## rename colums
data3 <- data2 %>%
  setNames(c('band', date_list,'dateUTC', 'station',
             'latitude','longitude','oxmgl','depth'))
names(data3)
```


### Data Perpare (simplified code)
```{r}
path_data <- './Img2Table/Img2Table_04_20191115'
sensor    <- 'aqua_'
# sensor    <- 'terr_'
yr        <- 2015

# read csv
pattern_file <- paste0('^', yr, '_', sensor, '*'); print(pattern_file)
csv_ls <- list.files(path = path_data, pattern = pattern_file, full.names = T)
csv_ls; length(csv_ls) ## should be 14

## a blank df
data      <- data.frame()
date_list <- seq(from = as.Date(paste0(yr, '-02-01')), 
                 to   = as.Date(paste0(yr, '-09-30')), by = 1) %>% as.character()

for (csv in csv_ls) {
  # print(csv)
  band_name <- gsub(pattern = '.csv', replacement = '', 
                    x = gsub(pattern = paste0(path_data, '/'), 
                             replacement = '', x = csv))
  print(band_name)
  csv_df <- read.csv(file = csv, header = T, stringsAsFactors = F)
  ## add band name col
  csv_df <- cbind(band = band_name, csv_df) 
  ## rename cols with date
  names(csv_df)[3:(3+length(date_list)-1)] <- as.character(date_list)
  ## select cols of interest
  csv_df <- csv_df %>% 
    dplyr::select(band, date_list, 
                  matches('date|station|latitude|longitude|oxmgl|depth', ignore.case = T)) %>%
    rename_all(tolower)  %>% 
    select(band, date_list, 
           'dateutc', 'station', 'latitude','longitude','oxmgl', everything()) %>%
    setNames(c('band', date_list,'dateutc', 'station', 'latitude','longitude','oxmgl','depth'))
  
  ## row bind
  data <- rbind(data, csv_df)
}



### save as csv
fname <- paste0('./Img2Table/data_processed/', 
                yr, '_', sensor, today, '.csv'); fname
# fname <- './Img2Table_2015_auqa/data_processed/terr_data_2015_update.csv'
# write.csv(x = data, file = fname)

names(data)
```


*data cleaning Test*
```{r include=FALSE}
### change all "No data" as NA
data2 <- data  %>% 
  mutate_all(~na_if(., 'No data'))

#### TEST: check UTC DATE foramt, and remove Special characters
####  for example, some date looks like 15-JUN-16, instead of 15JUN16
d1 = data2$dateutc[1]; d1

## 1) date include special char, such as "-"
d2 = gsub("[^0-9A-Za-z///' ]","" , d1 ,ignore.case = TRUE); d2
nchar(d2)

## 2) year using 2 or 4 digits
d3 = as.Date(d2, "%d%b%y"); d3 # %y	Year (2 digit)
d3 = as.Date(d2, "%d%b%Y"); d3 # %Y	Year (4 digit)

## 3) date in milliseconds, instead of normal date format 
##    https://developers.google.com/earth-engine/tutorial_js_02
### define a convert function 
ms_to_date = function(ms, t0="1970-01-01", timezone) {
        ## @ms: a numeric vector of milliseconds (big integers of 13 digits)
        ## @t0: a string of the format "yyyy-mm-dd", specifying the date that
        ##      corresponds to 0 millisecond
        ## @timezone: a string specifying a timezone that can be recognized by R
        ## return: a POSIXct vector representing calendar dates and times        
        sec = as.numeric(ms) / 1000
        as.POSIXct(sec, origin=t0, tz=timezone)
}

d31 <- ms_to_date(d2, timezone="UTC") %>% as.Date()
d31 
# as.Date(d5, format = "%Y-%m-%d %H:%M:%S")
# as.Date(d5)
nchar(d2)
str(d2)

d4 <- case_when(nchar(d2) <= 7  ~ as.Date(d2, "%d%b%y"),
                nchar(d2) <  13 ~ as.Date(d2, "%d%b%Y"), ## len = 9
                nchar(d2) >= 13 ~ as.Date(ms_to_date(d2, timezone="UTC")));
d4
```

### Data cleaning
```{r}
### change all "No data" as NA
data2 <- data  %>% 
  mutate_all(~na_if(., 'No data'))

names(data2)
### wide to long format
data3 <- data2 %>% as.data.frame() %>%
  gather(key = 'date_rs', value = 'rs', 2:(2+length(date_list)-1)) %>%
  dplyr::mutate(
    date_rs = as.Date(date_rs),
    DateUTC = gsub("[^0-9A-Za-z///' ]", "", x = dateutc, ignore.case = TRUE), 
              ### remove Special characters
    date_sp = case_when(nchar(DateUTC) >= 13 ~ as.Date(ms_to_date(DateUTC, timezone="UTC")),
                        nchar(DateUTC) <  13 ~ as.Date(DateUTC, "%d%b%Y"),
                        nchar(DateUTC) <= 7  ~ as.Date(DateUTC, "%d%b%y")),
    band    = gsub(pattern = paste0(yr, '_', sensor), replacement = '', x = band))

names(csv_df)
length(date_list)            ## how many days?     - 242
length(csv_df$band)  ## how many stations? - 
length(csv_ls)               ## how many bands?    - 14
length(date_list) * length(csv_df$band) * length(csv_ls) ## total obs.
length(date_list) * length(csv_df$band) * length(csv_ls) == length(data3[,1])

```




##   Viz
###  Window 1=8
#### Rrs_byBand_byDO in dif time window
```{r}
### filter a time_window
i <- 8*7 ## 8 weeks ago
t.window.i <- paste0(i, '_', i+7, '_d before'); t.window.i
w.window.i <- paste0('-', i/7+1, ' week'); t.window.i

names(data3)
data4 <- data3 %>%
  mutate(rs = as.numeric(rs))  %>%
  select('station', 'date_sp', 'oxmgl','date_rs', 'band', 'rs') %>%
  filter(date_rs <= (date_sp - i) & date_rs >= (date_sp - i -7)) %>%
  mutate(t.window = paste0(i, '_', i+7, '_d before'),
         w.window = paste0('-', i/7+1, ' week'))



### change to the similar scale of the numbers
data5 <- data4 %>%
  dplyr::filter(!is.na(rs)) %>%
  dplyr::mutate(
    rs2 = case_when(band == 'chlor_a' ~ rs*2,
                    band == 'nflh'    ~ rs*10^2,
                    band == 'poc'     ~ rs/10^2,
                    band == 'sst'     ~ rs,
                    TRUE              ~ rs*10^4)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(station, oxmgl, band, t.window, w.window) %>%
  dplyr::summarise(Rrs_mean = mean(rs2, na.rm = TRUE)) %>% as.data.frame()

  
names(data5)
## same way, but if_else look more complex
# data5 <- data4 %>%
#   filter(!is.na(rs)) %>%
#   dplyr::mutate(
#     rs2 = if_else(
#       band == 'chlor_a',   rs*2, if_else(
#         band == 'nflh',    rs*10^2, if_else(
#           band == 'poc',   rs/10^1/3, if_else(
#             band == 'sst', rs, rs*10^4))))) %>%
#   group_by(featureID, OxMgL, band) %>%
#   summarise(Rrs_mean = mean(rs2, na.rm = TRUE))

hist(data5$Rrs_mean)




Plot_Rrs_byBand_byDO <- function(df) {
  ggplot(data = df, aes(x = band, y = (Rrs_mean), 
                        group = station,
                        color = oxmgl)) +
  geom_line(alpha = 0.5) +
  # geom_point(aes(color = OxMgL)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  xlab('Bands') + ylab('Rrs') 
}

### run function
Plot_Rrs_byBand_byDO(data5) + ggtitle(w.window.i)
fname <- paste0('./Img2Table/Figures/', 
                yr, '_', sensor, 'Rrs_byBand_byDO_', 
                t.window.i, '1.png'); fname
ggsave(filename = fname, plot = last_plot(), 
       dpi = 300, width = 7, height = 4)

# data6 <- data5 %>%
#   filter(band == 'chlor_a')

hist(data5$Rrs_mean)
min(data5$Rrs_mean)

x= data5$Rrs_mean
n <- length(x)
sort(x, partial = 2)[1:50]
```




#### Rrs_vs_DO
```{r}
Plot_Rrs_vs_DO <- function(df) {
  ggplot(data = df, aes(y = oxmgl, 
                           # x = Rrs_mean,
                           x = log(Rrs_mean),
                           color = oxmgl)) +
    geom_point(alpha = 0.8) +
    geom_smooth(method = 'loess', formula = 'y ~ x') +
    facet_wrap(~band, scales = 'free_x')+
    geom_hline(yintercept=2, linetype="dashed", color = "gray", size=0.5)+
    theme_bw()+ 
    ggtitle(w.window.i) +
    ylab('DO (mg/l)') 
}

### run function
Plot_Rrs_vs_DO(data5)
fname <- paste0('./Img2Table/Figures/', 
                yr, '_', sensor, 'Rrs_vs_OD_', 
                t.window.i, '2.png'); fname
ggsave(filename = fname, plot = last_plot(), dpi = 300, width = 7, height = 6)
```



###  Loop by time.window
#### Data
```{r}
### filter time window
### select images with a week of sampling
seq(0, 63, by = 7)
dfs <- data.frame()
for (i in seq(0, 63, by = 7)) {
  print(- (i/7+1))
  # print(- i - 7)
  print(paste0(i, '_', i+7, '_d before'))
  print('-----------------------------')
  
  data41 <- data3 %>%
    dplyr::mutate(rs = as.numeric(rs))  %>%
    dplyr::select('station', 'date_sp', 'oxmgl', 'date_rs', 'band', 'rs') %>%
    # filter(featureID == 'id_1') %>%
    dplyr::filter(date_rs <= (date_sp -  i) & date_rs >= (date_sp - i - 7)) %>%
    dplyr::mutate(t.window = paste0(i, '_', i+7, '_d before'),
                  w.window = paste0('-', i/7+1, ' week'))

  # data5 <- data4 %>%
  #   filter(!is.na(rs)) %>%
  #   dplyr::mutate(
  #     rs2 = if_else(
  #       band == 'chlor_a',   rs, if_else(
  #         band == 'nflh',    rs*10^2, if_else(
  #           band == 'poc',   rs/10^2, if_else(
  #             band == 'sst', rs, rs*10^4))))) %>%
  #   group_by(featureID, OxMgL, band, t.window) %>%
  #   summarise(Rrs_mean = mean(rs2, na.rm = TRUE)) %>% as.data.frame()
  
  data51 <- data41 %>%
    dplyr::filter(!is.na(rs)) %>%
    dplyr::mutate(
      rs2 = case_when(band == 'chlor_a' ~ rs*2,
                      band == 'nflh'    ~ rs*10^2,
                      band == 'poc'     ~ rs/10^2,
                      band == 'sst'     ~ rs,
                      TRUE              ~ rs*10^4)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(station, oxmgl, band, t.window, w.window) %>%
    dplyr::summarise(Rrs_mean = mean(rs2, na.rm = TRUE)) %>% as.data.frame()

  dfs <- rbind(dfs, data51)
}

unique(dfs$t.window)
```


#### Rrs_byBand_byDO
```{r}
# ggplot(data = dfs, aes(x = band, y = (Rrs_mean), 
#                        group=featureID,
#                        color = OxMgL)) +
#   geom_line(alpha = 0.4, size = 1) +
#   # geom_point(aes(color = OxMgL)) +
#   facet_wrap(~t.window)+
#   theme_bw()+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))

Plot_Rrs_byBand_byDO(dfs) + facet_wrap(~w.window)
fname <- paste0('./Img2Table/Figures/', 
                yr, '_', sensor, 'Rrs_byBnad_byDO_byTimeWindow2.png'); 
fname
ggsave(filename = fname, plot = last_plot(), dpi = 200, width = 14, height = 7)

colSums(is.na(dfs))
df.na <- dfs %>% filter(is.na(Rrs_mean))
```



#### Rrs_vs_DO
```{r}
for (i in seq(0, 63, by = 7)) {
  t.window.i = paste0(i, '_', i+7, '_d before'); print(t.window.i)
  w.window.i = paste0('-', i/7+1, ' week'); print(w.window.i)
  dfs.i <- dfs %>%
    dplyr::filter(w.window == w.window.i)
  
  ## plot
  Plot_Rrs_vs_DO(dfs.i)
  fname <- paste0('./Img2Table/Figures/', yr, '_', sensor,
                  'Rrs_vs_OD_',
                  w.window.i, '.png'); fname
  ggsave(filename = fname, plot = last_plot(),
         dpi = 300, width = 7, height = 6)
}
```







###############################################
#  Pool multiple yrs together and plot
###############################################

## Data
```{r}
path_data <- './Img2Table/Img2Table_04_20191115'
# sensor    <- 'aqua_'
sensor    <- 'terr_'
# year      <- 2015

# read csv
pattern_file <- paste0('^201', '(.*)', sensor, '(.*)'); print(pattern_file)

csv_ls <- list.files(path = path_data, pattern = pattern_file, full.names = T)
csv_ls; length(csv_ls) ## should be 14

(2018-2010 + 1)*14

## a blank df
data      <- data.frame()
date_df   <- data.frame()

for (csv in csv_ls) {
  print(csv)
  yr <- as.numeric(substr(csv, 35, 38)); print(yr)
  date_list <- seq(from = as.Date(paste0(yr, '-02-01')),
                   to   = as.Date(paste0(yr, '-09-30')), by = 1) %>% as.character()
  n_days <- length(date_list)
  y_d    <- data.frame(year = yr, n_d = n_days)
  date_df <- rbind(date_df, y_d)
}

date_df <- unique(date_df)
## to account the total number of days across years
n_days_yrs <- 0

for (csv in csv_ls) {
  print(csv)
  ## read in csv
  csv_df <- read.csv(file = csv, header = T, stringsAsFactors = F)
  
  ### get the year info, and date list
  yr         <- as.numeric(substr(csv, 35, 38)); print(yr)
  ## way #1
  # date_list  <- seq(from = as.Date(paste0(yr, '-02-01')),to   = as.Date(paste0(yr, '-09-30')), by = 1) %>% as.character()
  ## way #2
  ##  for 2016 terra data, 2016-02-18 ~ 02-26 are missing, so way #1 will cause errors
  csv_dates  <- csv_df %>% dplyr::select(matches(paste0('X', yr), ignore.case = T)); names(csv_dates)
  date_list  <- names(csv_dates) %>% gsub('X', '', .) %>% gsub('\\.', '-', .); date_list
  
  n_days     <- length(date_list); print(n_days)
  n_days_yrs <- n_days_yrs + n_days

  ## add band name col
  ## to be added as new col
  band_name  <- gsub(pattern = '.csv', replacement = '',
                    x = gsub(pattern = paste0(path_data, '/'),
                             replacement = '', x = csv)); print(band_name)
  csv_df1    <- cbind(band = band_name, csv_df)
  
  ## rename cols with date
  names(csv_df1)[3:(3+length(date_list)-1)] <- as.character(date_list)
  ## select cols of interest
  csv_df2 <- csv_df1 %>%
    dplyr::rename_all(tolower)  %>%
    dplyr::select(band, date_list,
                  matches('dateutc|station|latitude|longitude|oxmgl|depth', ignore.case = T)) %>%
    dplyr::select(band, date_list,
           'dateutc', 'station', 'latitude','longitude','oxmgl', everything()) %>%
    setNames(c('band', date_list,
               'dateutc', 'station', 'latitude','longitude','oxmgl','depth')) %>%
    gather(key = date_rs, value = rs, ... = 2:(2+length(date_list)-1))

  ## row bind
  data <- rbind(data, csv_df2)
}
```

## - test
```{r eval=FALSE, include=FALSE}

csv <- "./Img2Table/Img2Table_04_20191115/2016_terr_chlor_a.csv"

yr <- as.numeric(substr(csv, 35, 38)); print(yr)
date_list <- seq(from = as.Date(paste0(yr, '-02-01')),
                 to   = as.Date(paste0(yr, '-09-30')), by = 1) %>% as.character()
n_days <- length(date_list); print(n_days)
n_days_yrs <- n_days_yrs + n_days

## to be added as new col
band_name <- gsub(pattern = '.csv', replacement = '',
                  x = gsub(pattern = paste0(path_data, '/'),
                           replacement = '', x = csv))
print(band_name)
  
  
## read in csv
csv_df    <- read.csv(file = csv, header = T, stringsAsFactors = F); names(csv_df)
csv_dates <- csv_df %>% dplyr::select(matches(paste0('X', yr), ignore.case = T)); names(csv_dates)
date_list <- names(csv_dates) %>% gsub('X', '', .) %>% gsub('\\.', '-', .); date_list


## add band name col
csv_df1 <- cbind(band = band_name, csv_df) ; names(csv_df1)
## rename cols with date
names(csv_df1)[3:(3+length(date_list)-1)] <- as.character(date_list); names(csv_df1)
## select cols of interest
csv_df2 <- csv_df1 %>% 
  dplyr::rename_all(tolower)  %>% 
  dplyr::select(band, date_list, 
                matches('dateutc|station|latitude|longitude|oxmgl|depth', ignore.case = T))# %>%
  dplyr::select(band, date_list, 
         'dateutc', 'station', 'latitude','longitude','oxmgl', everything()) %>%
  setNames(c('band', date_list,
             'dateutc', 'station', 'latitude','longitude','oxmgl','depth')) %>%
  gather(key = date_rs, value = rs, ... = 2:(2+length(date_list)-1))
  
names(csv_df1)
```


```{r}
## check the number of data record
length(data$band) ## 27510
days_ls <- as.vector(date_df$n_d);       days_ls
st_n_yrs<- st_n_yrs %>% filter(year>=2010) 
st_ls <- as.vector(st_n_yrs$count_st); st_ls
st_ls*14*days_ls

## check data number
sum(st_ls*14*days_ls)
nrow(data)
head(data)

names(data)
unique(data$dateutc)

## check data
## 2018, 133 stations
d <- data %>%
  filter(date_rs > as.Date('2018-01-01'))
head(d)

unique(d$station); 
length(unique(d$station)) ## 133 in 2018
unique(d$band)
length(unique(d$band))    ## 14
length(unique(d$date_rs)) ## 242 in 2018 

133*14*242
nrow(d)

```



## Data cleaning
```{r}
### change all "No data" as NA
###
data2 <- data  %>% 
  mutate_all(~na_if(., 'No data')) %>%
  mutate(DateUTC = gsub("-", "", ## "[^0-9A-Za-z///' ]" ## remove Special characters
                        dateutc, ignore.case = TRUE), 
         date_st = case_when(nchar(DateUTC) >= 11 ~ as.Date(ms_to_date(DateUTC, timezone="UTC")),
                             nchar(DateUTC) <  11 ~ as.Date(DateUTC, "%d%b%Y"),
                             nchar(DateUTC) <= 7  ~ as.Date(DateUTC, "%d%b%y"))) %>%
  as.data.frame()

head(data2)
str(data2)
data.na <- data2 %>% filter(is.na(date_st))
head(data.na)
```


# Time window (1w) for 2015
```{r}
year       <- 2015
date_start <- as.Date(paste0(year, '-01-01')); date_start
date_end   <- as.Date(paste0(year, '-12-31')); date_end
                 
data15 <- data2 %>%
  filter(date_st <= date_end & date_rs >= date_start) 

# ## test code
# i <- 1 ## time window, start from i weeks before sampling date
# data15_1w <- data15 %>% 
#   mutate(rs = as.numeric(rs))  %>%
#   filter(date_rs <= (date_st - 7*(i-1)) & date_rs > (date_st - 7*i)) %>%
#   mutate(t.window = paste0('1w_', i)) %>%
#   select(-dateutc, -DateUTC) %>%
#   group_by(station,latitude,longitude,oxmgl,depth, ## only keep station is ok
#            band, t.window) %>%
#   summarise(Rrs_mean = mean(rs, na.rm = TRUE)) %>% as.data.frame()



## loop for each time window
## create a week list
week_ls <- seq(1,8,1); week_ls
data15_1w_windows <- data.frame()
for (i in week_ls) {
  print(i)
  data15_1w <- data15 %>% 
    mutate(rs = as.numeric(rs))  %>%
    filter(date_rs <= (date_st - 7*(i-1)) & date_rs > (date_st - 7*i)) %>%
    mutate(t.window = paste0('1w_', i)) %>%
    select(-dateutc, -DateUTC) %>%
    group_by(station,latitude,longitude,oxmgl,depth, ## only keep station is ok
             band, t.window) %>%
    summarise(Rrs_mean = mean(rs, na.rm = TRUE)) %>% as.data.frame()
  ## row bind all
  data15_1w_windows <- rbind(data15_1w_windows, data15_1w)
}

## wide format
data15_1w_windows_wide <- data15_1w_windows %>%
  spread(key = band, value = Rrs_mean)

## Saving as RData 
getwd()
fname <- paste0('./Img2Table/data_', year, '_', sensor, '_by1w_long.RData'); fname
save(data15_1w_windows,     file = fname)
fname <- paste0('./Img2Table/data_', year, '_', sensor, '_by1w_wide.RData'); fname
save(data15_1w_windows_wide,file = fname )



ggplot(data = data15_1w_windows,
       aes(x=t.window, y=(Rrs_mean))) + 
  geom_boxplot() + 
  facet_wrap(~band, scales = 'free_y') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

getwd()
pname <- paste0('./Img2Table/Figures/', 'boxplot_', year, '_', sensor, '_by1w.png')
pname
ggsave(filename = pname, plot = last_plot(), width = 14, height = 10, dpi = 200)

```
